### 2020-02-21 4-Object-detection.tex

\chapter{Object detection}
Based on what is going to be detected, maybe text, cars, people?