### 2020-02-26 8-Conclusions.tex

\chapter{Conclusions}
Do some final review on why multi-task learning is a good idea given the problems shown and some personal experiments and what are the problems it brings.
Show how a multi-task setting is a trade-off between inference time and training time difficulty to motivate when it is a good idea or not.