### 2020-04-01 2-Background.tex
Sync

\chapter{Neural networks}
Basic background stuff, maybe cover little if there is room otherwise just add a reference to some basic material
\section{Fully Connected Networks}
When needed example of a transfer learning head?, when good/bad
\section{Convolutional Neural Networks}
Show some basic CNN things maybe and why needed over fully connected bois :)